---
title: "2019 Electronic Sales Exploratory Data Analysis"
author: "Kevin Banh"
date: "2/22/2022"
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

We'll be looking at sales data for electronics for a store during the entirety of 2019. Some exploratory data analysis will be done to gain some insights on the data.

The source of the data is https://github.com/KeithGalli/Pandas-Data-Science-Tasks/tree/master/SalesAnalysis/Sales_Data


# Libraries

```{r}
# Data visualization
library(ggplot2)
library(RColorBrewer)
library(ggtext)
library(scales)

# Data manipulation
library(tidyverse)
library(stringr)
library(reshape)
```


# Data processing

## Merging 12 months of data

```{r}
months <- c("January", "February", "March", "April", "May", "June", 
            "July", "August", "September", "October", "November", "December")

# Create an empty dataframe so we can append each month's data to
df <- data.frame()

for(i in months){
  # Create a string for the data name of each month
  data_name <- paste0("Sales_", i,"_2019.csv")
  curr_df <- read.csv(data_name)
  
  # Append to our dataframe
  df <- rbind(df, curr_df)
}
```

## Data cleaning

```{r}
glimpse(df)
```

Some things to note - Price.Each should be a double variable, Quantity.Ordered should be a double, and we can benefit from separating the date and times of Order.Date.

```{r}
df <- df %>%
  mutate(Month = substr(Order.Date, 1, 2), # Extract month from date
         Month = as.integer(Month), # Convert to integer
         Price.Each = as.double(Price.Each), 
         Quantity.Ordered = as.integer(Quantity.Ordered),
         Time = str_sub(Order.Date, -5, -4),
         Time = as.integer(Time),
         Order.ID = as.integer(Order.ID),
         Sales = Price.Each*Quantity.Ordered, 
         State = str_sub(Purchase.Address, -8, -7) # Get state portion of address
          ) 

glimpse(df)
```

Looks good! Now, let's try to handle any duplicate records.

```{r}
# Find duplicate records
df %>%
  group_by_all() %>%
  filter(n() > 1) %>%
  ungroup()

# There's 1428 duplicate rows so let's just keep the distinct rows
df <- df %>%
  distinct()

# Check for consistency how many rows are removed in case we messed up somewhere
dim(df)

# Let's test again to see if it worked
df %>% 
  group_by_all() %>%
  filter(n() > 1) %>%
  ungroup()
```

We get an empty table meaning there are no more duplicate records.


Let's tackle any null values that are not spread across a whole record

```{r}
sum(is.na(df))
```

```{r}
for(i in 1:dim(df)[2]){
  count = 0
  for(j in 1:dim(df)[1]){
    if(is.na(df[j,i])){
      count = count + 1
    }
  }
  print(paste0(names(df)[i], ": ", count))
}
```

Since we have **very few** NA values, we'll just delete them entirely especially since we cannot perform any kind of imputation on 
the fields that contain them.

```{r}
df <- na.omit(df)
```


# Exploratory date analysis

```{r}
plot_settings <- function(x){
  # x is a ggplot object
  
  x <- x + 
      theme(legend.position = "none", 
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_blank(),
      panel.background = element_blank(),
      axis.title.x = element_text(hjust = 0),
      axis.title.y = element_text(hjust = 1),
      axis.title = element_text(color = "#808080"),
      plot.title = element_markdown(size = 17)) + 
      scale_y_continuous(labels = comma) + 
      scale_fill_manual(values = c("royalblue4", "#808080"))  + 
      scale_fill_manual(values = c("royalblue4", "#808080"))
  return(x)
}
```


```{r}
g <- df %>%
  ggplot(aes(x = factor(Month), y = Sales, fill = factor(ifelse(Month == "12", "highlighted", "normal")))) + 
  geom_bar(stat = "identity") + 
  xlab("Month") + 
  labs(title = "<b style='color:#808080'>Sales was at its
    <b style='color:#27408b'>highest
    <b style='color:#808080'>during
    <b style='color:#27408b'>December</b>") + 
  scale_fill_manual(values = c("royalblue4", "#808080"))

plot_settings(g)
```




```{r}
g2 <- df %>%
  group_by(Product) %>%
  summarize(Quantity.Ordered = sum(Quantity.Ordered)) %>%
  arrange(desc(Quantity.Ordered)) %>%
  top_n(10) %>%
  ggplot(aes(x = factor(Product, levels = Product), y = Quantity.Ordered, 
             fill = ifelse(Product == "AAA Batteries (4-pack)", "highlighted",
                           ifelse(Product == "AA Batteries (4-pack)", "highlighted",
                                  "normal")))) +
    geom_bar(stat = "identity") + 
    ylab("Quantity Ordered") + 
    xlab("Product") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    labs(title = "<b style='color:#27408b'>Batteries 
         <b style='color:#808080'>are the top two most bought products</b>") +
  scale_fill_manual(values = c("royalblue4", "#808080"))

plot_settings(g2)



```



```{r}
g3 <- df %>%
  group_by(Product, Price.Each) %>%
  summarize(Quantity.Ordered = sum(Quantity.Ordered)) %>%
  arrange(desc(Quantity.Ordered)) %>%
  head(10) %>%
  ggplot(aes(x = factor(Product, levels = Product), y = Price.Each, 
             fill = ifelse(Product == "AAA Batteries (4-pack)", "highlighted",
                           ifelse(Product == "AA Batteries (4-pack)", "highlighted",
                                  "normal")))) +
    geom_bar(stat = "identity") + 
    ylab("Price") + 
    xlab("Product") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    labs(title = "<b style='color:#808080'>The most bought items are among the 
         <b style='color:#27408b'>cheapest</b>") +
  scale_fill_manual(values = c("royalblue4", "#808080"))

plot_settings(g3)
```



```{r}
g4 <- df %>%
  ggplot(aes(x = State, y = Quantity.Ordered,
             fill = ifelse(State == "CA", "highlighted", "normal"))) + 
  geom_bar(stat = "Identity") + 
  ylab("Quantity Ordered") + 
  labs(title = "<b style='color:#27408b'>California
       <b style='color:#808080'>had the most products ordered</b>") + 
  scale_fill_manual(values = c("royalblue4", "#808080"))
plot_settings(g4)
```


```{r}
g5 <- df %>%
  group_by(Time) %>%
  summarize(Quantity.Ordered = sum(Quantity.Ordered)) %>%
  ggplot(aes(x = factor(Time, levels = Time), y = Quantity.Ordered)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) + 
  geom_line(aes(group = 1), color = "#808080", size = 1.5) + 
  xlab("Time") + 
  ylab("Quantity Ordered") + 
  labs(title = "<b style='color:#27408b'>Noon 
       <b style='color:#808080'>and 
       <b style='color:#27408b'>evening 
       <b style='color:#808080'>times are peak ordering times</b>")
plot_settings(g5)
```

# Conclusion

Through our EDA, we've been able to answer the following questions

1. What part of the day had peak number of products sold?
2. What state ordered the most products?
3. What were the most popular items?
4. What were the costs of those popular items?
5. Which part of the month had the highest number of sales?

Knowing the answers to these will help the business better target their consumers. For example, they should try focusing on a wide-range of products that have a low price and are useful to a wide range of audiences. AAA/AA batteries were the most bought items and were among the cheapest as well. Furthermore, they could target their California consumers more since they make up a vast chunk of the products bought. 